---
title: "Tooth Growth on Guinea Pigs"
author: "mickfenneck"
output: pdf_document
---



Introduction
========

The project is meant to answer to this particular question:

> Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 

> 1. Load the ToothGrowth data and perform some basic exploratory data analyses 
> 2. Provide a basic summary of the data.
> 3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose. (Use the techniques from class even if there's other approaches worth considering)
> 4. State your conclusions and the assumptions needed for your conclusions. 

The Solution
========
### Analysis

We first have to give a look at the dataset and describe what is contained. Online you can find a quick description of the dataset:

Description
The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

Format
A data frame with 60 observations on 3 variables.

[,1]     len	 numeric	 Tooth length

[,2]	 supp	 factor	 Supplement type (VC or OJ).

[,3]	 dose	 numeric	 Dose in milligrams.

```{r}
set.seed(1994)
data(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)
head(ToothGrowth)
```

We have a continuous dependent variable [len] and two independent variables, dose and suppliment, that are factors with 3 and 2 different treatment levels,

```{r, echo=FALSE, fig.height = 4}
par(mfrow=c(1,2))
boxplot(len~supp, data=ToothGrowth, 
  col = c("yellow", "orange"),
  main = "Tooth Growth", 
  xlab = "Suppliment",
  ylab = "Tooth length")

boxplot(len~dose, data=ToothGrowth, 
  col = c("yellow", "orange", "cyan"),
  main = "Tooth Growth", 
  xlab = "Dose",
  ylab = "Tooth length")
```

```{r, echo=FALSE, fig.height = 4}
par(mfrow=c(1,1))
boxplot(len ~ dose, data = ToothGrowth,
        boxwex = 0.25, at = 1:3 - 0.2,
        subset = supp == "VC", col = "yellow",
        main = "Guinea Pigs' Tooth Growth",
        xlab = "Vitamin C dose mg",
        ylab = "tooth length",
        xlim = c(0.5, 3.5), ylim = c(0, 35), yaxs = "i")
boxplot(len ~ dose, data = ToothGrowth, add = TRUE,
        boxwex = 0.25, at = 1:3 + 0.2,
        subset = supp == "OJ", col = "orange")
legend(2, 9, c("Ascorbic acid", "Orange juice"),
       fill = c("yellow", "orange"))
```

### Comparing Dosage Levels
Using the t.test function given by default we can display the t statistic values

``` {r dosage level tests, echo=FALSE}
vc05 <- ToothGrowth[1:10, 1]
vc1 <- ToothGrowth[11:20, 1]
vc2 <- ToothGrowth[21:30, 1]
oj05 <- ToothGrowth[31:40, 1]
oj1 <- ToothGrowth[41:50, 1]
oj2 <- ToothGrowth[51:60, 1]

vc1vs05 <- t.test(vc1, vc05)
vc2vs1 <- t.test(vc2, vc1)
oj1vs05 <- t.test(oj1, oj05)
oj2vs1 <- t.test(oj2, oj1)

dosageLevelTable <- 
      data.frame(round(c(vc1vs05$statistic, vc2vs1$statistic, oj1vs05$statistic, oj2vs1$statistic), 3),
                 round(c(vc1vs05$parameter, vc2vs1$parameter, oj1vs05$parameter, oj2vs1$parameter), 3),
                 c(vc1vs05$p.value, vc2vs1$p.value, oj1vs05$p.value, oj2vs1$p.value),
                 c(toString(round(vc1vs05$conf.int, 3)), 
                   toString(round(vc2vs1$conf.int, 3)), 
                   toString(round(oj1vs05$conf.int, 3)), 
                   toString(round(oj2vs1$conf.int, 3))),
                 c(toString(round(vc1vs05$estimate, 3)),
                   toString(round(vc2vs1$estimate, 3)),
                   toString(round(oj1vs05$estimate, 3)),
                   toString(round(oj2vs1$estimate, 3))))
                
                 
colnames(dosageLevelTable) <- c("t", "df", "p-value", "conf interval", "sample estimates")
rownames(dosageLevelTable) <- c("VC 1.0 vs 0.5:  ", "VC 2.0 vs 1.0:  ", "OJ 1.0 vs 0.5:  ", "OJ 2.0 vs 1.0:  ")

dosageLevelTable


```

The table explains, as we should imagine, that ther is a correlation between the assumption of a major dosage of vitamine C and the growth of the teeth.

### Conclusion
We can finally state that the somministration of vitamine C in guinea pigs can determine an increase of the teeth growth in the animal.